---
title: "CSS Seed Predator"
author: "Jenna Loesberg"
date: "3/16/2021"
output:
  pdf_document: default
  html_document: default
---

Data analysis for CSS paper:

```{r, warning=FALSE,echo=FALSE, include = FALSE, message = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gt))
suppressPackageStartupMessages(library(webshot))

theme_set(theme_classic())
seed <- read.csv("C:/Users/Jenna/Dropbox/Jenna/seed predator/CSS_SeedPredator/CSS_seed_removal.csv") #this has major problem sites taken out
species <- read.csv("C:/Users/Jenna/Dropbox/Jenna/seed predator/CSS_SeedPredator/CSS_granivore_spp.csv") # without broken camera sites

#creating info to add to final data
info <- data.frame(species = c("salvia", "encelia", "brassica", "bromus"),
                   size = c("Small", "Large", "Small", "Large"),
                   origin = c("Native", "Native", "Invasive","Invasive"))
```

Which sites to drop:

- sites where more than 5 seeds were missing from Exclusion dish
- sites where less than 10 or more than 70 seeds taken
- sites where seeds taken are less than 10 different from Exclusion

```{r}
# Selecting exclusion dishes with 5 or less seeds missing, there are 94
exc <- seed %>% 
   filter(Treatment == "C")  #101 observations
  

exc <- exc %>%
  filter(Total_eaten <=5) #94 observations

# get unique ID for those sites:
exc_codes <- exc %>% 
  select(GPS_code)
# add back the rest of the data
seed <- semi_join(seed, exc_codes) #seed = without those Exc sites

# BR and A only dishes with seed eaten no. between 10 and 70
BR <- seed %>% 
  filter(Treatment == "BR+") %>% #88 dishes
  filter(Total_eaten  %in% (10:70)) #52 dishes

A <- seed %>% 
  filter(Treatment == "A+") %>% #90 dishes
  filter(Total_eaten  %in% (10:70)) #5 dishes
ABR <- rbind(BR, A)
# ABR: These are the sites we're keeping.

# do any of these sites have <10 differnece with control?
GPS <- ABR %>% 
  select(GPS_code) #these are the sites that we need control numbers for 

GPS <- semi_join(exc, GPS) #control sites for each site that meet criteria

ABR <- rbind(GPS, ABR)
ABR2 <- ABR
#one site has a too small seed difference between C and trt:

ABR <- ABR %>% 
  filter(GPS_code != "JL_Feb15_S2") %>% 
  arrange(GPS_code, Treatment)


# These are the sites that do not have camera data: let's exclude them from the analyses

ABR <- ABR %>% 
  filter(GPS_code != "May22_S2") %>% 
  filter(GPS_code != "Nov5_S4") %>% 
  filter(GPS_code != "Oct10_S1") %>% 
  filter(GPS_code != "Sep17_S4") %>% 
  filter(GPS_code != "Ap23_S2") %>% 
  filter(GPS_code != "Ap23_S4")

#ABR has all of the sites that we're including! Now getting just A and BR
Global_final <- ABR %>% 
  filter(Treatment != "C")
```


Creating histograms of seeds missing for appendix (before trials are cleaned for preference analysis):

```{r}
#Changing label names for fact wrap
treatment_names <- c("BR+" = "Bird and Rodent",
                     "A+" = "Ant",
                     "C" = "Exclusion")

seed %>% 
  ggplot(aes(x = Total_eaten, fill = Treatment))+
  geom_histogram(binwidth=5, color = "black")+
  facet_wrap(~Treatment, labeller = as_labeller(treatment_names))+
  scale_fill_grey(start=0.6, end=0.2) +
  theme(legend.position = "none")+
  labs(x = "Total number of seeds removed", y = "Number of dishes")

#ggsave("figures/histogram.png", width = 8, height = 4)
```

creating Manly alpha indices:
```{r}
Global_final["prey"] <- 20 #starting amount  = 20
Global_final <- Global_final %>% 
  mutate(salv.pro = salvia_left/prey, #proportion of starting amount
         enc.pro = encelia_left/prey,
         brom.pro = bromus_left/prey,
         bras.pro = brassica_left/prey)
#add small vlaue to zeros to take log:
Global_final$salv.pro[Global_final$salv.pro == 0] <- 0.001
Global_final$enc.pro[Global_final$enc.pro == 0] <- 0.001
Global_final$brom.pro[Global_final$brom.pro == 0] <- 0.001
Global_final$bras.pro[Global_final$bras.pro == 0] <- 0.001

Global_final <- Global_final %>% 
  mutate(salvia = log(salv.pro)/(log(salv.pro)+log(enc.pro)+log(brom.pro)+log(bras.pro)), #manly calc
         encelia = log(enc.pro)/(log(salv.pro)+log(enc.pro)+log(brom.pro)+log(bras.pro)),
         bromus = log(brom.pro)/(log(salv.pro)+log(enc.pro)+log(brom.pro)+log(bras.pro)),
         brassica = log(bras.pro)/(log(salv.pro)+log(enc.pro)+log(brom.pro)+log(bras.pro)))

# splitting into A, BR and global

A_final <- Global_final %>% 
  filter(Treatment == "A+")
BR_final <- Global_final %>% 
  filter(Treatment == "BR+")
```

This is in wide format, need to switch to long

```{r}
BR_final <- BR_final %>%
  select(GPS_code, salvia, encelia, bromus, brassica) %>% 
  pivot_longer(cols = c(salvia:brassica), names_to = "species", values_to = "index")
A_final <- A_final %>% 
  select(GPS_code, salvia, encelia, bromus, brassica) %>% 
  pivot_longer(cols = c(salvia:brassica), names_to = "species", values_to = "index")
Global_final <- Global_final %>% 
  select(Treatment, GPS_code, salvia, encelia, bromus, brassica) %>% 
  pivot_longer(cols = c(salvia:brassica), names_to = "species", values_to = "index")
```

T-tests for each seed species. Are the true means different from 0.25?
```{r}
bromus <- Global_final %>% 
  filter(species == "bromus") %>% 
  select(index)
t.test(bromus, mu = 0.25)
brassica <- Global_final %>% 
  filter(species == "brassica") %>% 
  select(index)
t.test(brassica, mu = 0.25)
salvia <- Global_final %>% 
  filter(species == "salvia") %>% 
  select(index)
t.test(salvia, mu = 0.25)
encelia <- Global_final %>% 
  filter(species == "encelia") %>% 
  select(index)
t.test(encelia, mu = 0.25)
```


Confidence intervals for graph:
```{r}
BR_sum <- BR_final %>% 
  group_by(species) %>% 
  summarize(mean = mean(index), sd = sd(index), n = n(), se = sd/sqrt(n))
Global_sum <- Global_final %>% 
  group_by(species) %>% 
  summarize(mean = mean(index), 
            sd = sd(index), 
            n = n(), 
            se = sd/sqrt(n),
            lower.ci = mean-1.96*sd/sqrt(n),
            upper.ci = mean+1.96*sd/sqrt(n))


A_sum <- A_final %>% 
  group_by(species) %>% 
  summarize(mean = mean(index), sd = sd(index), n = n(), se = sd/sqrt(n))

Global_sum <- left_join(Global_sum, info, by = "species")
Global_sum$species[Global_sum$species == "brassica"] <- "Brassica"
Global_sum$species[Global_sum$species == "bromus"] <- "Bromus"
Global_sum$species[Global_sum$species == "encelia"] <- "Encelia"
Global_sum$species[Global_sum$species == "salvia"] <- "Salvia"

```

Global plot:
```{r}
Global_sum %>% 
  mutate(species = as.factor(species),
         size = as.factor(size)) %>% 
  ggplot(aes(x = factor(species, levels = c("Brassica","Bromus","Salvia", "Encelia")), y = mean))+
  geom_col(aes(fill = size), width = .95, color = "black")+
  #geom_point(data = Global_final, aes(species, index))+
  scale_fill_manual(values=c("darkslategrey", "gray88"), guide = guide_legend(reverse=TRUE))+
  geom_hline(aes(linetype = "Preference Threshold"), yintercept=0.25,linetype="dashed", color = "black", size = .8)+
  #scale_linetype_manual(name = "", values = c(1, 1), guide = guide_legend(override.aes = "black"))+
  geom_errorbar(aes(ymin = lower.ci, ymax = upper.ci), width = 0.2)+
  expand_limits(y = c(0, 1))+
  labs(y= "Manly-\u03B1 Preference Index", x = "", fill = "Seed Size")+
  facet_grid(.~origin, scales = "free", switch = "x", space = "free")+
   theme(strip.placement = "outside", 
         legend.position = c(0.87, 0.87),
         axis.text.x = element_text(face = "italic"),
         strip.text.x = element_text(size = 11))
#ggsave("figures/GlobalManly.png", width = 7, height = 5)

# n = 50
```

```{r, include= F}
A_sum %>% 
  ggplot(aes(species, mean))+
  geom_col()+
  geom_hline(yintercept=0.25,linetype="dashed", color = "red")+
  geom_errorbar(aes(ymin = mean - se, ymax = mean+se))+
  expand_limits(y = c(0, 1))+
  labs(y= "Manly Index", title = "Ant preference (n = 5)")
#ggsave("ant_manly_Aug18.png", width = 10, height = 7)

BR_sum %>% 
  ggplot(aes(species, mean))+
  geom_col()+
  geom_hline(yintercept=0.25,linetype="dashed", color = "red")+
  geom_errorbar(aes(ymin = mean - se, ymax = mean+se))+
  expand_limits(y = c(0, 1))+
  labs(y= "Manly Index", title = "BR preference (n = 50)")
#ggsave("BR_manly_Aug18.png", width = 10, height = 7)
```

Add in the granivore species information from the camera:

```{r}
site_species <- species %>% 
  select(c(Run, GPS_code, Site, Sp_ate))

site_species <- left_join(Global_final, site_species, by = "GPS_code")
site_species$Sp_ate <- ifelse(site_species$Treatment == "A+", "ant", site_species$Sp_ate)
site_species <- left_join(site_species, info)
```

species specific plot:


```{r}
labels <- c(ant = "Ant", CA_Towee = "California Towhee", GS = "Ground Squirrel", mixed_bird = "Multiple Birds", mixed_mammal = "Multiple Mammals", multipred = "Multiple Guilds", none = "No Visits", scrub_jay = "Scrub Jay", WC_sparrow = "WC Sparrow", woodrat = "Woodrat")



site_species <- site_species %>% 
  filter(Sp_ate != "none") %>%
  mutate(Sp_ate = as.factor(Sp_ate),
         species = as.factor(species),
         species = factor(species, levels = c("brassica","bromus","salvia", "encelia")),
         Sp_ate = factor(Sp_ate, levels = c("mixed_bird", "CA_Towee", "scrub_jay", "WC_sparrow", "ant", "multipred", "mixed_mammal", "woodrat", "GS")))

species_stat <- site_species %>% 
  group_by(species, Sp_ate) %>% 
  summarize(mean = mean(index)) %>% 
  filter(Sp_ate == "mixed_bird"| Sp_ate =="CA_Towee"|Sp_ate =="ant"| Sp_ate =="multipred" )

site_species %>% 
  group_by(species) %>% 
  ggplot(aes(species, index))+
  geom_point(aes(fill = size), shape = 21, size = 1.8)+
  geom_hline(yintercept=0.25,linetype="dashed", color = "black", size = .5)+
  expand_limits(y = c(0, 1))+
  facet_wrap(~Sp_ate, labeller = labeller(Sp_ate = labels), ncol = 3)+
  labs(y= "Manly-\u03B1 Preference Index", x = "", fill = "Seed Size")+
  theme(axis.text.x=element_text(angle=50, hjust=1)) +
  geom_point(data = species_stat, aes(y = mean, x = species), shape = 95, size = 7, color = "#000000")+
  scale_x_discrete(labels = c("Brassica","Bromus","Salvia", "Encelia"))+ #just changes names
  theme(axis.text.x = element_text(face = "italic"))+
  scale_fill_manual(values=c("darkslategrey", "gray88"), guide = guide_legend(reverse=TRUE))
#ggsave("figures/specices_preference_1.png", width = 7, height = 6 )
  
```

One sample t-tests for all of those means:

- Multiple birds:
```{r}
mult_birds_seeds <- site_species %>% 
  filter(Sp_ate == "mixed_bird")

MBbird_bromus <- mult_birds_seeds %>% 
  filter(species == "bromus") %>% 
  select(index)
t.test(MBbird_bromus, mu = 0.25)

MBbird_brassica <- mult_birds_seeds %>% 
  filter(species == "brassica") %>% 
  select(index)
t.test(MBbird_brassica, mu = 0.25)

MBbird_salvia <- mult_birds_seeds %>% 
  filter(species == "salvia") %>% 
  select(index)
t.test(MBbird_salvia, mu = 0.25)

MBbird_encelia <- mult_birds_seeds %>% 
  filter(species == "encelia") %>% 
  select(index)
t.test(MBbird_encelia, mu = 0.25)
```

- towhee:
```{r}
towhee_seeds <- site_species %>% 
  filter(Sp_ate == "CA_Towee")

towhee_bromus <- towhee_seeds %>% 
  filter(species == "bromus") %>% 
  select(index)
t.test(towhee_bromus, mu = 0.25)

towhee_brassica <- towhee_seeds %>% 
  filter(species == "brassica") %>% 
  select(index)
t.test(towhee_brassica, mu = 0.25)

towhee_salvia <- towhee_seeds %>% 
  filter(species == "salvia") %>% 
  select(index)
t.test(towhee_salvia, mu = 0.25)

towhee_encelia <- towhee_seeds %>% 
  filter(species == "encelia") %>% 
  select(index)
t.test(towhee_encelia, mu = 0.25)
```

- ant:
```{r}
ant_seeds <- site_species %>% 
  filter(Sp_ate == "ant")

ant_bromus <- ant_seeds %>% 
  filter(species == "bromus") %>% 
  select(index)
t.test(ant_bromus, mu = 0.25)

ant_brassica <- ant_seeds %>% 
  filter(species == "brassica") %>% 
  select(index)
t.test(ant_brassica, mu = 0.25)

ant_salvia <- ant_seeds %>% 
  filter(species == "salvia") %>% 
  select(index)
t.test(ant_salvia, mu = 0.25)

ant_encelia <- ant_seeds %>% 
  filter(species == "encelia") %>% 
  select(index)
t.test(ant_encelia, mu = 0.25)
```
- Multiple taxa:
```{r}
mult_taxa_seeds <- site_species %>% 
  filter(Sp_ate == "multipred")

mult_taxa_bromus <- mult_taxa_seeds %>% 
  filter(species == "bromus") %>% 
  select(index)
t.test(mult_taxa_bromus, mu = 0.25)

mult_taxa_brassica <- mult_taxa_seeds %>% 
  filter(species == "brassica") %>% 
  select(index)
t.test(mult_taxa_brassica, mu = 0.25)

mult_taxa_salvia <- mult_taxa_seeds %>% 
  filter(species == "salvia") %>% 
  select(index)
t.test(mult_taxa_salvia, mu = 0.25)

mult_taxa_encelia <- mult_taxa_seeds %>% 
  filter(species == "encelia") %>% 
  select(index)
t.test(mult_taxa_encelia, mu = 0.25)
```

Now onto the community table...

```{r}
# take out non-granivore species
species2 <- species[-c(5, 18, 21, 26, 34, 35, 36)]
  
```

Changing the dataframe into a useable format:

```{r}

species[species == 0] <- NA

species <- species %>% 
  summarize_if(is.numeric, sum, na.rm=TRUE) # add up all of the observations
  
species <- species %>% 
  pivot_longer(cols = -c(Site), values_to = "number" , names_to = c("species", "type"), names_sep = "_") 
species[is.na(species)] <- 0
  
species <- species %>% 
  pivot_wider(names_from = "type", values_from = "number") %>% 
  select(-c(Site)) 
species[is.na(species)] <- 0
species <- species%>% 
  group_by(species) %>% 
  mutate(ratio = Eat/(Visit+Eat))
            
```


